### 2021-05-18 paper/paper.Rmd
Create JOSS paper folder

---
# Example from https://joss.readthedocs.io/en/latest/submitting.html
title: 'Latentcor: An R Package for Latent Correlation Estimation'
tags:
  - R
  - Statistics
  - Latent Correlation
authors:
  - name: Mingze Huang
    orcid: 0000-0003-3919-1564
    affiliation: "1, 2" # (Multiple affiliations must be quoted)
  - name: Irina Gaynanova
    orcid: 0000-0000-0000-0000
    affiliation: 2
  - name: Christian L. Muller
affiliations:
 - name: Department of Statistics, Texas A& M University
   index: 1
 - name: Department of Economics, Texas A& M University
   index: 2
 - name: Department of Statistics, University of Munich
   index: 3
date: 18 May 2021
year: 2021
bibliography: paper.bib
output: rticles::joss_article
csl: apa.csl
journal: JOSS
---

# Summary

# Mathematics

# Citations

# Rendered R Figures

# Acknowledgements

# References