### 2021-05-31 paper/paper.Rmd
Create reference table.

---
title: 'Latentcor: An R Package for Latent Correlation Estimation'
tags:
- R
- Statistics
- Latent Correlation
date: "18 May 2021"
output:
  rticles::joss_article: default
  rticles::joss_article(): default
authors:
- name: Mingze Huang
  orcid: 0000-0003-3919-1564
  affiliation: 1, 2
- name: Irina Gaynanova
  orcid: 0000-0000-0000-0000
  affiliation: 2
- name: Christian L. Muller
year: 2021
bibliography: paper.bib
affiliations:
- name: Department of Statistics, Texas A& M University
  index: 1
- name: Department of Economics, Texas A& M University
  index: 2
- name: Department of Statistics, University of Munich
  index: 3
csl: apa.csl
journal: JOSS
---

# Summary

The R package *latentcor* provides estimation for latent correlation with mixed data types (continuous, binary, truncated and ternary). Comparing to *MixedCCA*, which estimates latent correlation for canonical correlation analysis, our new package provides a standalone version for latent correlation estimation. Also we add new functionality for latent correlation between ternary/continous, ternary/binary, ternary/truncated and ternary/ternary cases.

Compare to MixedCCA, standalone, new functionality, memory footprint.

# Statement of need
Currently there is no standalone package dealing with latent correlation for mixed data type like we did in *latentcor*. The R package *stats* [@team2013r] have some functionality to calculate different type of correlations (Pearson, Kendall and Spearman). The R package *pcaPP* [@croux2013robust] provides a fast calculation for Kendall's \tau. The R package *MixedCCA* [@yoon2020sparse] have functionality for latent correlation estimation as an intermediate step for canonical correlation analysis on mixed data.
No package deal with latent correlation across mixed data type.

# Usage

 |Type | continuous | binary | truncated | ternary |
|-----|----------|
|continuous | @liu2009nonparanormal | @fan2017high | @yoon2020sparse | @quan2018rank |
|binary | @fan2017high | @fan2017high | @yoon2020sparse | @quan2018rank |
|truncated | @yoon2020sparse | @yoon2020sparse | @yoon2020sparse | This paper |
|ternary | @quan2018rank | @quan2018rank | This paper | @quan2018rank |

 Definition for Kendall tau.
 Describe the theorem states expected value of kendall tau = F(r). refer to table for reference of formula.
 A sentence to describe two methods: slow original vs interpolation.
 Table to show memory improvement compare to mixedCCA.
 Generate data for pairs.
 One example of original call, one for interpolation call.

# Rendered R Figures

# References